# (APPENDIX) Appendix {-}

# R Code Reference {#coderef}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = 'hold')
# knitr::opts_chunk$set(class.source = 'code-style')
# knitr::opts_chunk$set(class.output = 'out-style')
set.seed(12222)
```

## Binomial distribution

For the following examples, I am using the following labeled values so you can see the kind of result R will give you.

```{r}
this_prob <- 0.3
this_size <- 10

given_value <- 5

upper_value <- 9
lower_value <- 3

given_probability <- 0.2
```

To use these bits of code yourself, just replace the label with the appropriate number or assign a new value under the same label (e.g., `this_prob <- 0.941`).

### Probability of observing an outcome exactly equal to a given value

```{r}
dbinom(x = given_value, prob = this_prob, size = this_size)
```

### Probability of observing a value less than or equal to a given value

```{r}
pbinom(q = given_value, prob = this_prob, size = this_size)
```

### Probability of observing a value greater than a given value

```{r}
1 - pbinom(q = given_value, prob = this_prob, size = this_size)
```

### Probability of observing a value inside an interval between an upper and lower value

```{r}
pbinom(q = upper_value, prob = this_prob, size = this_size) - pbinom(q = lower_value, prob = this_prob, size = this_size)
```

### Probability of observing a value outside an interval between an upper and lower value

```{r}
1 - (pbinom(q = upper_value, prob = this_prob, size = this_size) - pbinom(q = lower_value, prob = this_prob, size = this_size))
```

or

```{r}
1 - pbinom(q = upper_value, prob = this_prob, size = this_size) + pbinom(q = lower_value, prob = this_prob, size = this_size)
```

### The value for which there is a given probability of observing an outcome less than or equal to that value

```{r}
qbinom(p = given_probability, prob = this_prob, size = this_size)
```

### Testing a null hypothesis

The following are important bits of code for conducting null hypothesis tests with the binomial distribution.  For the following examples, I have told R to remember the following labeled values, in addition to the ones above.

```{r}
null_prob <- 0.5

sample_value <- 6

alpha_level <- 0.1
```

#### One-tailed test (alternative is that probability is *less* than null probability)

##### Boundary of confidence interval

```{r}
ci_boundary <- qbinom(p = alpha_level, prob = null_prob, size = this_size)
```

##### $p$ value

```{r}
pbinom(q = sample_value, prob = null_prob, size = this_size)
```

#### One-tailed test (alternative is that probability is *greater* than null probability)

##### Boundary of confidence interval

```{r}
ci_boundary <- qbinom(p = 1 - alpha_level, prob = null_prob, size = this_size)
```

##### $p$ value

```{r}
1 - pbinom(q = sample_value - 1, prob = null_prob, size = this_size)
```

## Normal distribution

For the following examples, I am using the following labeled values so you can see the kind of result R will give you.

```{r}
this_mean <- 0
this_sd <- 1

given_value <- 1.5

upper_value <- 2
lower_value <- -1

given_probability <- 0.2
```

To use these bits of code yourself, just replace the label with the appropriate number or assign a new value under the same label (e.g., `this_mean <- -4`).

### Probability of observing a value less than or equal to a given value

```{r}
pnorm(q = given_value, mean = this_mean, sd = this_sd)
```

### Probability of observing a value greater than a given value

```{r}
1 - pnorm(q = given_value, mean = this_mean, sd = this_sd)
```

### Probability of observing a value inside an interval between an upper and lower value

```{r}
pnorm(q = upper_value, mean = this_mean, sd = this_sd) - pnorm(q = lower_value, mean = this_mean, sd = this_sd)
```

### Probability of observing a value outside an interval between an upper and lower value

```{r}
1 - (pnorm(q = upper_value, mean = this_mean, sd = this_sd) - pnorm(q = lower_value, mean = this_mean, sd = this_sd))
```

or

```{r}
1 - pnorm(q = upper_value, mean = this_mean, sd = this_sd) + pnorm(q = lower_value, mean = this_mean, sd = this_sd)
```

### The value for which there is a given probability of observing an outcome less than or equal to that value

```{r}
qnorm(p = given_probability, mean = this_mean, sd = this_sd)
```

### Testing a null hypothesis

The following are important bits of code for conducting null hypothesis tests with the normal distribution (see [this lab](#lab8) for additional examples).  For the following examples, I have told R to remember the following labeled values.

```{r}
null_mean <- 0
null_sd <- 1

sample_size <- 30

sample_mean <- 0.3

effect_size <- 0.5

alpha_level <- 0.1
```

#### Standard error of the mean

```{r}
null_sd / sqrt(sample_size)
```

#### One-tailed test (alternative is that mean is *less* than null mean)

##### Boundary of confidence interval

```{r}
ci_boundary <- qnorm(p = alpha_level, mean = null_mean, sd = null_sd / sqrt(sample_size))
```

##### $p$ value

```{r}
pnorm(q = sample_mean, mean = null_mean, sd = null_sd / sqrt(sample_size))
```

##### Power

```{r}
pnorm(q = ci_boundary, mean = null_mean + effect_size, sd = null_sd / sqrt(sample_size))
```

#### One-tailed test (alternative is that mean is *greater* than null mean)

##### Boundary of confidence interval

```{r}
ci_boundary <- qnorm(p = 1 - alpha_level, mean = null_mean, sd = null_sd / sqrt(sample_size))
```

##### $p$ value

```{r}
1 - pnorm(q = sample_mean, mean = null_mean, sd = null_sd / sqrt(sample_size))
```

##### Power

```{r}
1 - pnorm(q = ci_boundary, mean = null_mean + effect_size, sd = null_sd / sqrt(sample_size))
```

#### Two-tailed test

##### Boundaries of confidence interval

```{r}
ci_lower <- qnorm(p = alpha_level / 2, mean = null_mean, sd = null_sd / sqrt(sample_size))
ci_upper <- qnorm(p = 1 - alpha_level / 2, mean = null_mean, sd = null_sd / sqrt(sample_size))
```

##### $p$ value

If `sample_mean > null_mean`:

```{r}
2 * (1 - pnorm(q = sample_mean, mean = null_mean, sd = null_sd / sqrt(sample_size)))
```

If `sample_mean <= null_mean`:

```{r echo=FALSE}
sample_mean <- -abs(sample_mean)
```

```{r}
2 * pnorm(q = sample_mean, mean = null_mean, sd = null_sd / sqrt(sample_size))
```

```{r echo=FALSE}
sample_mean <- abs(sample_mean)
```

##### Power

```{r}
pnorm(q = ci_lower, mean = null_mean + effect_size, sd = null_sd / sqrt(sample_size)) +
    1 - pnorm(q = ci_upper, mean = null_mean + effect_size, sd = null_sd / sqrt(sample_size))
```